Does pre-crastination explain why some observers are sub-optimal in a visual search task?

### 2023-09-29 RegReport/paper.tex
future tense removed down to results section

We will recruit 200 participants for this experiment. This large sample size is due to the addition of personality measures: Larger samples provide greater power to detect the typical effect sizes in those fields, reduce Type I error, and increase the chances of replicable findings \cite{fraley2014}. Participants will be recruited via participant panels and opportunity sampling at the universities of Aberdeen and Essex. The protocol has been approved by the ethics committees at both universities. Participants will complete the bucket-retrieval task both before and after the visual search task, so no counterbalancing will be needed. Additional participants will be recruited to replace any who are excluded from data analysis (details below).

@@

Figure \ref{fig:buckets} is a photograph of the bucket task set-up. Two 10 litre buckets will be filled with sand until the total weight of each bucket is 3kg. The experiment will take place in a quiet corridor. Buckets will be placed at 6ft and 16ft away from the start position, which is marked as a T with green tape. Buckets will be placed on crates at about hand height. Just behind the start position is a pair of crates on which to set the buckets down to end the trial.

@@

The task replicates that used in \cite{fournier2019task}, in which participants were instructed to pick up two buckets and place them on the crates behind the start position. We will closely follow the methods and instructions as described in their appendix. Each participant will complete three trials before the visual search task, and three trials after. While the experimenter is setting up the buckets for the next trial, the participant will be asked to walk over to a nearby window and look outside until they are asked to return. The full protocol and instructions given to participants is available at https://osf.io/5aq4c/.

@@

Participants will be classified as "pre-crastinators" if they pick up the closer bucket first, and "optimal" if they pick up the far bucket first. Based on the pilot data, and that of \cite{fournier2019task}, we expect the results from the bucket task to be largely binary, with participants picking up either the close or the far bucket on every trial. If the data are similarly binary in this experiment, we will continue to treat pre-crastination as a two-level categorical factor. Participants who pick up the closest bucket on at least five of the six trials will be classed as pre-crastinators, while those who pick up the closest bucket on no more than two out of six trials will be non-pre-crastinators. Participants with intermediate behaviour will be excluded.

There is also a chance that, unlike \cite{fournier2019task}, our sample of participants give rise to a larger range of intermediate behaviours. If more than $25\%$ of our participants would be excluded under the coding scheme outlined above, then we will treat our pre-crastination measure as a three-level factor (if a clear classification is apparent). In this case, the same Bayesian beta regression model will be fit to the data.

@@

Stimuli will be presented on a 17 inch CRT monitor with a resolution of $1024 \times 768$. Stimulus generation, presentation and data collection are controlled by Matlab and psychophysics toolbox \cite{brainard1997,cornelissen2002} run on a mac pro computer. The position of the right eye will be recorded using a desktop-mounted EyeLink 1000 eye tracker (SR Research, Canada) sampling eye position at 1000Hz.

We will use the same arrays of line segments as in \cite{nowakowska2017} (see Figure \ref{fig:stimulus}). These arrays consist of 22 column and 16 rows on a uniform grey background. The target line is always tilted $45^{\circ}$ degrees to the right, while the mean distractor angle is perpendicular at $45^{\circ}$ degrees to the left. Search difficulty will be manipulated by sampling from either a narrow $30^{\circ}$ range of distractor line orientations (homogeneous) or a wide $106^{\circ}$ range (heterogeneous). Half of each search array consists of homogeneous line segments, while the other half is heterogeneous. Which side is heterogeneous will be randomly determined on each trial. We will use 96 search trials in total, half of which will contain a target. This number of trials was selected based on simulating the results of shorter experiments using the data from \cite{nowakowska2017} (see Figure \ref{fig:num_trials}). The target could be located in any of the possible locations apart from the middle four vertical columns.

@@

\subsubsection{Procedure} Before the start of the visual search experiment, participants will undergo a nine-point calibration sequence. Following this, participants will be instructed to report, as quickly as possible, whether a line tilted $45^{\circ}$ to the right was present or not, using the up (present) or down (absent) arrow key. Each search trial begins with a black fixation point (letter \texttt{x}) subtending $1.5\times 2.5$cm ($1.9^{\circ}\times 3.1^{\circ}$), presented at the centre of the computer screen. On the press of a space bar, the stimulus is displayed until the participant makes a response (or times out after 60 seconds). Auditory feedback in the form of a beep, and a 2s red screen, immediately follows incorrect key presses. A set of 8 practice trials precedes the experimental trials.

@@

Data from participants who fail to achieve at least $75\%$ accuracy for \textit{easy} and \textit{absent} conditions will be excluded as it suggests that they did not understand the task. Fixations during trials will be classified as falling on the homogeneous or heterogeneous half of the display, with fixations falling within the central four columns of line segments being excluded. Following \cite{nowakowska2017} we will summarise each participant's strategy as the proportion of fixations two through five falling on the heterogeneous side of the display during target absent trials. Trials with fewer than five fixations will be excluded, and we will only analyse data from participants who have at least 10 trials.

@@

We will administer three personality questionnaires. The first questionnaire is the second iteration of the Big Five Inventory (BFI-2 \cite{soto2017}). The BFI-2 measures five broad personality traits: extraversion (e.g., ``is outgoing, sociable''), agreeableness (e.g., ``has a forgiving nature''), conscientiousness (e.g., ``is efficient, gets things done''), negative emotionality (e.g., ``can be tense''), and open-mindedness (e.g., ``is fascinated by art, music, or literature''). Based on previous research, we suspect that precrastination will have the strongest relationship with conscientiousness. Those who are high in conscientiousness are organised, productive, and responsible, and may have an urge to pick up a bucket as soon as possible.

The second personality questionnaire is the Barratt Impulsiveness Scale (BIS-11 \cite{patton1995}). The BIS-11 measures overall impulsiveness, as well as three subtypes: attentional impulsiveness (``I don't `pay attention'''), motor impulsiveness (``I act on the spur of the moment''), and non-planning impulsiveness (``I say things without thinking''). Based on previous research, we suspect that neither impulsiveness nor its subscales will be related to precrastination.

@@

Personality questionnaires will be administered to participants at the end of the experiment to prevent them from influencing the behavioural measures. Questionnaires will be hosted on Qualtrics and participants will access the questionnaires using a laboratory computer.

Participants will be allowed to skip questions they do not wish to answer. For any one questionnaire, a response rate of less than $75\%$ of questions, or giving the same response for more than $75\%$ of all questions, will invalidate the data from that questionnaire. Participants will be replaced if they answer less than $75\%$ of the questions on the conscientiousness scale (the others are exploratory).

@@

The aim of this experiment is to determine if there is a relationship between pre-crastination and the sub-optimal behaviour we observe in the split-half visual search task. Following \cite{nowakowska2017, clarke2019}, each participant's strategy will be represented by the proportion of fixations two-through-five directed to the difficult half of the search area. We will use a Bayesian approach to estimate the extent to which performance in the pre-crastination bucket task is predictive of visual search strategy. Our approach to Bayesian modelling will follow the advice given in \cite{mcelreath2016}, with models fitted using R \cite{r2019} and Stan \cite{carpenter2017stan}.

As our independent variable is a proportion, we will fit a model using beta regression \cite{ferrari2004beta} with a logit link for for $\mu$ (mean) and a log link for $\phi$ (precision). We will use conservative priors: $\beta \sim\mathcal{N}(0, 0.50)$ and $\gamma \sim\mathcal{N}(0, 1)$ (where $\beta$ and $\gamma$ are the coefficients for $mu$ and $\gamma$ respectively). This means that we are assuming that the mean of each group is 0.50, with a $95\%$ HPDI $= [0.27, 0.73]$\footnote{For comparison, the mean strategy score for the participants in \cite{nowakowska2017} with $n=14$ is 0.58, with a range of $[0.10, 0.99]$. Similarly, \cite{clarke2019} ($n=57$) found a mean of 0.60 with a range of $[0.06, 0.98]$.}. Our prior for the difference between groups is $\beta \sim\mathcal{N}(0, 0.50)$. We have included pilot data (above) to demonstrate our planned analysis. We have previously used this analysis approach with this paradigm to explore the effect of motivation on visual search strategy \cite{james2019}.

We will also measure the relationship between the personality measures and tendency to pre-crastinate, using similar methods to those outlined above. If our measures of pre-crastination turn out to be mainly binary, then we will use Bayesian logistic regression to measure the relationship, otherwise, we will report an $R^2$ statistic.

\subsection{Outcome-neutral conditions} The ability to test the stated hypotheses depends on replicating both the pre-crastination behaviour and the variability in visual search efficiency observed in previous studies. Our pilot data achieve this replication. Given that we plan to repeat the same conditions with a larger sample, it is unlikely that we will not be able to test our hypotheses. Nonetheless, in our pilot results, the proportion of participants who were categorized as precrastinators (13/30) was smaller than in the most similar previous experiment (Fournier et al., \cite{fournier2019task}, 36/49). Some variation in the proportion can be expected, but a very high ($>90\%$) or very low ($<10\%$) prevalence of pre-crastination will suggest there is something in the way we set up or instructed the bucket task that provoked an unrepresentative behaviour in our participants, and will not be able to use these data to test our hypotheses. It is also critical that we find inter-participant variability in search strategy. Our pilot data replicate this variation as well, and it is difficult to imagine why variability would be restricted in our larger sample. Nonetheless, if the variability in our sample is less than 0.1 (= half the standard deviation of the pilot data (0.2)), we will not proceed to use these data to test our hypotheses.

@@

The main (pre-registered) hypothesis motivating this experiment is that participants who pre-crastinate will also be sub-optimal (in terms of making more early eye movements to the homogeneous half of the display) in the split-half visual search task. After carrying out our pilot experiment, it looks like there may be also be a difference in speed-accuracy trade-offs. If this (or any other differences in eye movements) appears to be an interesting effect after we collect our full dataset, we will carry out suitable exploratory analysis. We will also investigate the possibility of further links between the personality measures, pre-crastination and visual search.

>>>>

\textcolor{red}{We planned to recruit} 200 participants for this experiment. This large sample size is due to the addition of personality measures: Larger samples provide greater power to detect the typical effect sizes in those fields, reduce Type I error, and increase the chances of replicable findings \cite{fraley2014}. Participants \textcolor{red}{were} recruited via participant panels and opportunity sampling at the universities of Aberdeen and Essex. The protocol \textcolor{red}{was} approved by the ethics committees at both universities. Participants \textcolor{red}{completed} the bucket-retrieval task both before and after the visual search task, so no counterbalancing \textcolor{red}{was} needed. Additional participants \textcolor{red}{were} recruited to replace \textcolor{red}{those} excluded from data analysis (details below).

@@

Figure \ref{fig:buckets} is a photograph of the bucket task set-up. Two 10 litre buckets \textcolor{red}{were} filled with sand until the total weight of each bucket \textcolor{red}{was} 3kg. The experiment \textcolor{red}{took} place in a quiet corridor. Buckets \textcolor{red}{were} placed at 6ft and 16ft away from the start position, which \textcolor{red}{was} marked as a T with green tape. Buckets \textcolor{red}{were} placed on crates at about hand height. Just behind the start position \textcolor{red}{was} a pair of crates on which to set the buckets down to end the trial.

@@

The task replicates that used in \cite{fournier2019task}, in which participants were instructed to pick up two buckets and place them on the crates behind the start position. We closely \textcolor{red}{followed} the methods and instructions as described in their appendix. Each participant \textcolor{red}{completed} three trials before the visual search task, and three trials after. While the experimenter \textcolor{red}{was} setting up the buckets for the next trial, the participant \textcolor{red}{was} asked to walk over to a nearby window and look outside until asked to return. The full protocol and instructions given to participants is available at https://osf.io/5aq4c/.

@@

Participants \textcolor{red}{were} classified as "pre-crastinators" if they \textcolor{red}{picked} up the closer bucket first, and "optimal" if they \textcolor{red}{picked} up the far bucket first. Based on the pilot data, and that of \cite{fournier2019task}, we \textcolor{red}{expected} the results from the bucket task to be largely binary, with participants picking up either the close or the far bucket on every trial. \textcolor{red}{Our plan was the following:} If the data are similarly binary in this experiment, we will continue to treat pre-crastination as a two-level categorical factor. Participants who pick up the closest bucket on at least five of the six trials will be classed as pre-crastinators, while those who pick up the closest bucket on no more than two out of six trials will be non-pre-crastinators. Participants with intermediate behaviour will be excluded.

There \textcolor{red}{was} also a chance that, unlike \cite{fournier2019task}, our sample of participants \textcolor{red}{would} give rise to a larger range of intermediate behaviours. If more than $25\%$ of our participants would \textcolor{red}{have been} excluded under the coding scheme outlined above, then we \textcolor{red}{would have treated} our pre-crastination measure as a three-level factor (if a clear classification is apparent). In this case, the same Bayesian beta regression model \textcolor{red}{would have been} fit to the data.

@@

Stimuli \textcolor{red}{were} presented on a 17 inch CRT monitor with a resolution of $1024 \times 768$. Stimulus generation, presentation and data collection \textcolor{red}{were} controlled by Matlab and psychophysics toolbox \cite{brainard1997,cornelissen2002} run on a mac pro computer. The position of the right eye \textcolor{red}{was} recorded using a desktop-mounted EyeLink 1000 eye tracker (SR Research, Canada) sampling eye position at 1000Hz.

We \textcolor{red}{used} the same arrays of line segments as in \cite{nowakowska2017} (see Figure \ref{fig:stimulus}). These arrays consist of 22 column and 16 rows on a uniform grey background. The target line is always tilted $45^{\circ}$ degrees to the right, while the mean distractor angle is perpendicular at $45^{\circ}$ degrees to the left. Search difficulty \textcolor{red}{was} manipulated by sampling from either a narrow $30^{\circ}$ range of distractor line orientations (homogeneous) or a wide $106^{\circ}$ range (heterogeneous). Half of each search array consists of homogeneous line segments, while the other half is heterogeneous. Which side is heterogeneous \textcolor{red}{was} randomly determined on each trial. We \textcolor{red}{used} 96 search trials in total, half of which \textcolor{red}{contained} a target. This number of trials was selected based on simulating the results of shorter experiments using the data from \cite{nowakowska2017} (see Figure \ref{fig:num_trials}). The target could be located in any of the possible locations apart from the middle four vertical columns.

@@

\subsubsection{Procedure} Before the start of the visual search experiment, participants undergo a nine-point calibration sequence. Following this, participants \textcolor{red}{were} instructed to report, as quickly as possible, whether a line tilted $45^{\circ}$ to the right was present or not, using the up (present) or down (absent) arrow key. Each search trial begins with a black fixation point (letter \texttt{x}) subtending $1.5\times 2.5$cm ($1.9^{\circ}\times 3.1^{\circ}$), presented at the centre of the computer screen. On the press of a space bar, the stimulus is displayed until the participant makes a response (or times out after 60 seconds). Auditory feedback in the form of a beep, and a 2s red screen, immediately follows incorrect key presses. A set of 8 practice trials precedes the experimental trials.

@@

Data from participants who fail to achieve at least $75\%$ accuracy for \textit{easy} and \textit{absent} conditions \textcolor{red}{were} excluded as it suggests that they did not understand the task. Fixations during trials \textcolor{red}{were} classified as falling on the homogeneous or heterogeneous half of the display, with fixations falling within the central four columns of line segments being excluded. Following \cite{nowakowska2017} we \textcolor{red}{summarised} each participant's strategy as the proportion of fixations two through five falling on the heterogeneous side of the display during target absent trials. Trials with fewer than five fixations \textcolor{red}{were} excluded, and we only \textcolor{red}{analysed} data from participants who have at least 10 trials.

@@

We \textcolor{red}{administered} three personality questionnaires. The first questionnaire is the second iteration of the Big Five Inventory (BFI-2 \cite{soto2017}). The BFI-2 measures five broad personality traits: extraversion (e.g., ``is outgoing, sociable''), agreeableness (e.g., ``has a forgiving nature''), conscientiousness (e.g., ``is efficient, gets things done''), negative emotionality (e.g., ``can be tense''), and open-mindedness (e.g., ``is fascinated by art, music, or literature''). Based on previous research, we \textcolor{red}{suspected} that precrastination will have the strongest relationship with conscientiousness. Those who are high in conscientiousness are organised, productive, and responsible, and may have an urge to pick up a bucket as soon as possible.

The second personality questionnaire is the Barratt Impulsiveness Scale (BIS-11 \cite{patton1995}). The BIS-11 measures overall impulsiveness, as well as three subtypes: attentional impulsiveness (``I don't `pay attention'''), motor impulsiveness (``I act on the spur of the moment''), and non-planning impulsiveness (``I say things without thinking''). Based on previous research, we \textcolor{red}{suspected} that neither impulsiveness nor its subscales will be related to precrastination.

@@

Personality questionnaires \textcolor{red}{were} administered to participants at the end of the experiment to prevent them from influencing the behavioural measures. Questionnaires \textcolor{red}{were} hosted on Qualtrics and participants \textcolor{red}{accessed} the questionnaires using a laboratory computer.

Participants \textcolor{red}{were} allowed to skip questions they \textcolor{red}{did} not wish to answer. For any one questionnaire, a response rate of less than $75\%$ of questions, or giving the same response for more than $75\%$ of all questions, \textcolor{red}{invalidated} the data from that questionnaire. Participants \textcolor{red}{would be} replaced if they \textcolor{red}{answered} less than $75\%$ of the questions on the conscientiousness scale (the others are exploratory).

@@

The aim of this experiment is to determine if there is a relationship between pre-crastination and the sub-optimal behaviour we observe in the split-half visual search task. Following \cite{nowakowska2017, clarke2019}, each participant's strategy \textcolor{red}{was} represented by the proportion of fixations two-through-five directed to the difficult half of the search area. We \textcolor{red}{used} a Bayesian approach to estimate the extent to which performance in the pre-crastination bucket task is predictive of visual search strategy. Our approach to Bayesian modelling \textcolor{red}{followed} the advice given in \cite{mcelreath2016}, with models fitted using R \cite{r2019} and Stan \cite{carpenter2017stan}.

As our independent variable is a proportion, we fit a model using beta regression \cite{ferrari2004beta} with a logit link for for $\mu$ (mean) and a log link for $\phi$ (precision). We \textcolor{red}{used} conservative priors: $\beta \sim\mathcal{N}(0, 0.50)$ and $\gamma \sim\mathcal{N}(0, 1)$ (where $\beta$ and $\gamma$ are the coefficients for $mu$ and $\gamma$ respectively). This means that we are assuming that the mean of each group is 0.50, with a $95\%$ HPDI $= [0.27, 0.73]$\footnote{For comparison, the mean strategy score for the participants in \cite{nowakowska2017} with $n=14$ is 0.58, with a range of $[0.10, 0.99]$. Similarly, \cite{clarke2019} ($n=57$) found a mean of 0.60 with a range of $[0.06, 0.98]$.}. Our prior for the difference between groups is $\beta \sim\mathcal{N}(0, 0.50)$. We have included pilot data (above) to demonstrate our planned analysis. We have previously used this analysis approach with this paradigm to explore the effect of motivation on visual search strategy \cite{james2019}.

We also \textcolor{red}{measured} the relationship between the personality measures and tendency to pre-crastinate, using similar methods to those outlined above. If our measures of pre-crastination \textcolor{red}{turned} out to be mainly binary \textcolor{red}{which they did}, then we \textcolor{red}{planned to} use Bayesian logistic regression to measure the relationship. Otherwise, we \textcolor{red}{would have reported} an $R^2$ statistic.

\subsection{Outcome-neutral conditions} The ability to test the stated hypotheses depends on replicating both the pre-crastination behaviour and the variability in visual search efficiency observed in previous studies. Our pilot data achieve this replication. Given that we \textcolor{red}{planned} to repeat the same conditions with a larger sample, it \textcolor{red}{as} unlikely that we \textcolor{red}{would} not be able to test our hypotheses. Nonetheless, in our pilot results, the proportion of participants who were categorized as precrastinators (13/30) was smaller than in the most similar previous experiment (Fournier et al., \cite{fournier2019task}, 36/49). Some variation in the proportion can be expected, but a very high ($>90\%$) or very low ($<10\%$) prevalence of pre-crastination \textcolor{red}{would have suggested} there is something in the way we set up or instructed the bucket task that provoked an unrepresentative behaviour in our participants, and \textcolor{red}{would} not be able to use these data to test our hypotheses. It \textcolor{red}{was} also critical that we find inter-participant variability in search strategy. Our pilot data replicate this variation as well, and it \textcolor{red}{was} difficult to imagine why variability would be restricted in our larger sample. Nonetheless, if the variability in our sample \textcolor{red}{had been} less than 0.1 (= half the standard deviation of the pilot data (0.2)), we \textcolor{red}{would not have proceeded} to use these data to test our hypotheses.

@@

The main (pre-registered) hypothesis motivating this experiment is that participants who pre-crastinate will also be sub-optimal (in terms of making more early eye movements to the homogeneous half of the display) in the split-half visual search task. After carrying out our pilot experiment, it \textcolor{red}{looked} like there may be also be a difference in speed-accuracy trade-offs. If this (or any other differences in eye movements) \textcolor{red}{had appeared} to be an interesting effect after we \textcolor{red}{collected} our full dataset, we \textcolor{red}{would have carried} out suitable exploratory analysis. We also \textcolor{red}{investigated} the possibility of further links between the personality measures, pre-crastination and visual search.
